---
title: "Leaflet_Assignment_1_Henry"
output:
  html_document: default
  pdf_document: default
date: "2024-04-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("leaflet")
library("widgetframe")
library("webshot")
```

# [GETTING STARTED WITH LEAFLET]{.underline}

*Try to work through down this script, observing what happens in the plotting pane. There are three*

*initial example exercises, followed by a few questions and 5 tasks.*

*Review favorite backgrounds in: <https://leaflet-extras.github.io/leaflet-providers/preview/> beware that some need extra options specified*

*To install Leaflet package, run this command at your R prompt:*

```{r}
#install.packages("leaflet") 
#install.packages("htmlwidget")
```

*Activate the library*

```{r}
library(leaflet) 
library(htmlwidgets) # not essential, only needed for saving the map as .html
```

# [EXAMPLES]{.underline}

## Example 1: create a Leaflet map of Europe with addAwesomeMarkers()

Learn to create a map by plotting three point markers, called Robin, Jakub and Jannes. First, create labels for your points:

```{r}
popup = c("Robin", "Jakub", "Jannes")
```

Next, you create a Leaflet map with these basic steps: call the leaflet() function and pipe in some background maps ("tiles"), and then addMarker() function with longitude and latitude. Run the pipeline below to see the result:

```{r}
leaflet() %>% # create a map widget by calling the leaflet library 
  addProviderTiles("Esri.WorldPhysical") %>% # add Esri World Physical map tiles explicitly 
  addAwesomeMarkers(lng = c(-3, 23, 11), # add point markers specified with longitude for 3 points 
                    lat = c(52, 53, 49), # and latitude for 3 points (coordinates fall in EU) 
                    popup = popup) # specify labels for points, which will appear if you click on the point in the map
```

## Example 2: create a Leaflet map of Sydney with the setView() function

Note that there are no points or markers below, just background tiles. setView() function alone helps you focus and zoom the map in a particular area of the world.

Now we are in Sydney

```{r}
leaflet() %>% addTiles() %>% # add default OpenStreetMap map tiles 
  addProviderTiles("Esri.WorldImagery", # add custom Esri World Physical map tiles 
                   options = providerTileOptions(opacity=0.5)) %>% # make the Esri tile transparent 
  setView(lng = 151.005006, lat = -33.9767231, zoom = 10) # set the location of the map 
```

[**Question 1: What is the order of longitude and latitude in the setView() function?**]{.underline}

*A1: Longitude goes before lattitude, similar to how the x-axis goes before the y-axis when reffering graphs.*

Now let's go back to Europe again. When the map defined below renders, click the box in top right corner in your map Viewer to select between different background layers

```{r}
leaflet() %>% 
  addTiles() %>% 
  setView( 
    lng = 2.34, 
    lat = 48.85, 
    zoom = 5 ) %>% # let's use setView to navigate to our area 
  addProviderTiles("Esri.WorldPhysical", group = "Physical") %>% # add physical background 
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>% # add satellite image 
  addProviderTiles("MtbMap", group = "Geo") %>% # add geomorphic units map
  addLayersControl( # we are adding layers control to the maps 
    baseGroups = c("Geo","Aerial", "Physical"), 
    options = layersControlOptions(collapsed = T)) # replace T with F and back and run it
```

**Question 2: How does the map above change if you replace the T in the last line of code above with F?**

*A2: When "layersControlOptions(collapsed = T)" the options for manipulating the map-layers are only revealed as the cursor hovers over the icon in the top right, when "layersControlOptions(collapsed = F)" the options for manipulating the map layers are always revealed, regardless of the position of the cursor.*

## Example 3: SYDNEY HARBOUR DISPLAY WITH 11 LAYERS

Let's create a map with more background layers that allows interactive selection between multiple layers. This can be useful if you want to rendering historical maps illustrating change over time. The chunk below sets the location and zoom level of your map:

```{r}
leaflet() %>% 
  setView(151.2339084, 
          -33.85089, 
          zoom = 13) %>% 
  addTiles() # checking I am in the right area
```

Bring in a choice of Esri background layers, Create a basic base map

```{r}
l_aus <- leaflet() %>% # assign the base location to an object 
  setView(151.2339084, 
          -33.85089, 
          zoom = 13)
```

Now, prepare to select backgrounds by grabbing their names

```{r}
esri <- grep("^Esri", 
             providers, 
             value = TRUE)
```

Select backgrounds from among provider tiles. To view the options, go to <https://leaflet-extras.github.io/leaflet-providers/preview/>.

*Run the following three lines together!*

```{r}
for (provider in esri) { l_aus <- l_aus %>% 
  addProviderTiles(provider, 
                   group = provider) }
```

Map of Sydney, NSW, Australia. We make a layered map out of the components above and write it to an object called '*AUSmap*'

```{r}
AUSmap <- l_aus %>% 
  addLayersControl(baseGroups = names(esri), 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  addMiniMap(tiles = esri[[1]], 
             toggleDisplay = TRUE, 
             position = "bottomright") %>% 
  addMeasure( position = "bottomleft", 
              primaryLengthUnit = "meters", 
              primaryAreaUnit = "sqmeters", 
              activeColor = "#3D535D", 
              completedColor = "#7D4479") %>% 
  htmlwidgets::onRender(" function(el, x) { var myMap = this; myMap.on('baselayerchange', function (e) { myMap.minimap.changeLayer(L.tileLayer.provider(e.name)); }) }") %>% 
  addControl("", position = "topright")
```

*run this to see your product*

```{r}
AUSmap
```

#### SAVE THE FINAL PRODUCT

Save map as a html document (optional, replacement of pushing the export button) only works in root, we will also need this widget to make pretty maps:

```{r}
saveWidget(AUSmap, 
           "AUSmap.html", 
           selfcontained = TRUE)
```

# TASK NUMBER ONE:

*"Create a Danish equivalent of AUSmap with Esri layers, but call it DANmap. You will need it layer as a background for Danish data points."*

## Solution1\*:

\**using coordinates retrieved from 3rd party*

```{r}
leaflet() %>% setView(10.38831, 
                      55.39594, 
                      zoom = 6) %>% 
  addTiles()

l_den<- leaflet() %>% 
  setView(10.38831, 
          55.39594, 
          zoom = 6)

for (provider in esri) { l_den <- l_den %>% 
  addProviderTiles(provider, 
                   group = provider) }

DENmap <- l_den %>% 
  addLayersControl(baseGroups = names(esri), 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  addMiniMap(tiles = esri[[1]], 
             toggleDisplay = TRUE, 
             position = "bottomright") %>% 
  addMeasure( position = "bottomleft", 
              primaryLengthUnit = "meters", 
              primaryAreaUnit = "sqmeters", 
              activeColor = "#3D535D", 
              completedColor = "#7D4479") %>% 
  htmlwidgets::onRender(" function(el, x) { var myMap = this; myMap.on('baselayerchange', function (e) { myMap.minimap.changeLayer(L.tileLayer.provider(e.name)); }) }") %>% 
  addControl("", 
             position = "topright")

DENmap
saveWidget(DENmap, "DENmap.html", 
                  selfcontained = TRUE)
```

## ADD DATA TO LEAFLET

Before you can proceed to Task 2, you need to learn about coordinate creation. In this section you will manually create machine-readable spatial data from GoogleMaps, load these into R, and display them in Leaflet with addMarkers(): First, go to <https://bit.ly/CreateCoordinates1>. Enter the coordinates of your favorite leisure places in Denmark extracting them from the URL in googlemaps, adding name and type of monument. Remember to copy the coordinates as a string, as just two decimal numbers separated by comma.

### *Caveats: Do NOT edit the grey columns! They populate automatically!*

Second: read the sheet into R. You will need gmail login information.

IMPORTANT: watch the console, it may ask you to authenticate or put in the number that corresponds to the account you wish to use.

Needed libraries:

```{r}
library(tidyverse) 
library(googlesheets4) 
library(leaflet)
```

If you experience difficulty with your read_sheet() function (it is erroring out), uncomment and run the following function:

```{r}
gs4_deauth() 
```

run this line and then rerun the read_sheet() function below Read in the Google sheet you've edited

```{r}
places <- read_sheet("https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918", 
                     col_types = "cccnncnc", # check that you have the right number and type of columns 
                     range = "DM2023") # select the correct worksheet name

glimpse(places)
```

### [Question 3: are the Latitude and Longitude columns present?]{.underline}

*A3: yes*

### [Do they contain numeric decimal degrees?]{.underline}

*A3(2): yes*

If your coordinates look good, see how you can use addMarkers() function to load them in a basic map. Run the lines below and check: are any points missing? Why?

```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude,  
             lat = places$Latitude, 
             popup = paste(places$Description, 
                           "<br>", 
                           places$Type)) 
```

Now that you have learned how to load points from a googlesheet to a basic leaflet map, apply the know-how to YOUR DANmap object.

**Solution:**

```{r}
l_den<- leaflet() %>% 
  setView(lng = places$Longitude[120], 
          lat = places$Latitude[120], 
          zoom = 6) 

for (provider in esri) { l_den <- l_den %>% 
            addProviderTiles(provider, 
                             group = provider) }
DENmap <- l_den %>% 
  addLayersControl(baseGroups = names(esri), 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  addMiniMap(tiles = esri[[1]], 
             toggleDisplay = TRUE, 
             position = "bottomright") %>% 
  addMeasure( position = "bottomleft", 
              primaryLengthUnit = "meters", 
              primaryAreaUnit = "sqmeters", 
              activeColor = "#3D535D", 
              completedColor = "#7D4479") %>% 
  htmlwidgets::onRender(" function(el, x) { var myMap = this; myMap.on('baselayerchange', function (e) { myMap.minimap.changeLayer(L.tileLayer.provider(e.name)); }) }") %>% 
  addControl("", 
             position = "topright")
DENmap 
saveWidget(DENmap, "DENmap.html", 
           selfcontained = TRUE)
```

# TASK NUMBER TWO:

*"Read in the googlesheet data you and your colleagues created into your DANmap object (with 11 background layers you created in Task 1)."*

**Solution:**

```{r}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = places$Longitude,  
             lat = places$Latitude, 
             popup = paste(places$Description, 
                           "<br>", 
                           places$Type)) 
DENmap_popups<- l_den %>% 
  addLayersControl(baseGroups = names(esri), 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  addMiniMap(tiles = esri[[1]], 
             toggleDisplay = TRUE, 
             position = "bottomright") %>% 
  addMeasure( position = "bottomleft", 
              primaryLengthUnit = "meters", 
              primaryAreaUnit = "sqmeters", 
              activeColor = "#3D535D", 
              completedColor = "#7D4479") %>% 
  addMarkers(lng = places$Longitude,  
             lat = places$Latitude, 
             popup = paste(places$Description, 
                           "<br>", 
                           places$Type)) %>% 
  htmlwidgets::onRender(" function(el, x) { var myMap = this; myMap.on('baselayerchange', function (e) { myMap.minimap.changeLayer(L.tileLayer.provider(e.name)); }) }") %>% 
  addControl("", 
             position = "topright")

DENmap_popups 
saveWidget(DENmap_popups, "DENmap_popups.html", selfcontained = TRUE)
```

# TASK NUMBER THREE:

*"Can you cluster the points in Leaflet? Hint: Google "clustering options in Leaflet in R"*

**Solution:**

```{r}
DENmap_popups_cluster<- l_den %>% 
  addLayersControl(baseGroups = names(esri), 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  addMiniMap(tiles = esri[[1]], 
             toggleDisplay = TRUE, 
             position = "bottomright") %>% 
  addMeasure( position = "bottomleft", 
              primaryLengthUnit = "meters", 
              primaryAreaUnit = "sqmeters", 
              activeColor = "#3D535D", 
              completedColor = "#7D4479") %>% 
  addMarkers(lng = places$Longitude,  
             lat = places$Latitude, 
             popup = paste(places$Description, 
                           "<br>", places$Type), 
             clusterOptions = markerClusterOptions( )) %>% 
  htmlwidgets::onRender(" function(el, x) { var myMap = this; myMap.on('baselayerchange', function (e) { myMap.minimap.changeLayer(L.tileLayer.provider(e.name)); }) }") %>%
  addControl("", 
             position = "topright")

DENmap_popups_cluster 
saveWidget(DENmap_popups_cluster, "DENmap_popups_cluster.html", selfcontained = TRUE)
```

# TASK NUMBER FOUR:

*"Look at the two maps (with and without clustering) and consider what each is good for and what not."*

**Your brief answer:**

*the clustered version (DENmap_popups_cluster) preserves more of the geographic map beneath, the unclustered version (DENmap_popups) technically provides the user with the option to directly access the markers,in practice this is made difficult by the sheer number of markers, at least in this case. Another advantage of the unclustered version is that it kind of 'clusters' itself when multiple markers overlap, this however again comes at the expense of the map #beneath.*

# TASK NUMBER FIVE:

*Find out how to display the notes and classifications column in the map. Hint: Check online help in sites such as <https://r-charts.com/spatial/interactive-maps-leaflet/#popup>*

**Solution:**

*NOTE: i did not understand what was meant by 'Classification' so i added Coordinates instead, also i formatted the text, which i think turned out pretty cool:)*

```{r}
DENmap_popups_clusterplus<- l_den %>% 
  addLayersControl(baseGroups = names(esri), 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
                   addMiniMap(tiles = esri[[1]], 
                   toggleDisplay = TRUE, 
                   position = "bottomright") %>% 
                   addMeasure( position = "bottomleft", 
                     primaryLengthUnit = "meters", 
                     primaryAreaUnit = "sqmeters", 
                     activeColor = "#3D535D", 
                     completedColor = "#7D4479") %>% 
                   addMarkers(lng = places$Longitude,  
                    lat = places$Latitude, 
                    popup = paste("<strong>",
                                  "<i>",
                                  places$Description,
                                  "</i>", 
                                  "</strong>", 
                                  "<hr>", 
                                  "<strong>", 
                                  "Type:",
                                  "</strong>", 
                                  places$Type,
                                  "<br>",
                                  "<strong>", 
                                  "Notes:",
                                  "</strong>", 
                                  places$Notes[is.na(places$Notes)] <- paste("<i>", 
                                                                             "No notes have been added to this location", 
                                                                             "</i>"), 
                                  "<hr>", 
                                  "<i>", 
                                  places$CoordinatesFILLTHIS, 
                                  "</i>"), 
                    clusterOptions = markerClusterOptions()) %>% 
                    htmlwidgets::onRender(" function(el, x) { var myMap = this; myMap.on('baselayerchange', function (e) { myMap.minimap.changeLayer(L.tileLayer.provider(e.name)); }) }")

DENmap_popups_clusterplus 
saveWidget(DENmap_popups_clusterplus, "DENmap_popups_clusterplus.html", selfcontained = TRUE)
```

######################################## CONGRATULATIONS - YOUR ARE DONE :)

*I SO AM!:)*
